---
title: "“Hongxiao Yu Final Report”"
author: "“Hongxiao Yu”"
date: "“2 May 2020”"
output:
  html_document:
    df_print: paged
Output: “html_notebook”
---

***
### source function
```{r}
library(here)
source(here("Function", "function for analysis.R"))
```
***


***
### Part #l Data Preparation
*1.1 tidy the Job Prep Demographics*
*1.1.1Import and tidy the Job Prep Demographics file into a tibble named “demo”.*
```{r}
# note that I have transfer the discipline column on Excel.
library(here)
library(tidyverse)
demo <- read_csv(here("Data","Job Prep Demographics2.csv"),
                 skip = 5, 
                 col_names = TRUE,
                 col_types = list(ID = col_double(), 
                                  "School Code" = col_character(), 
                                  Age = col_character(),
                                  Gender = col_character()))
   
demo
```


*1.1.2 Eliminate rows with missing data*
```{r}
demo <- na.omit(demo)
demo
```


*1.1.3 A primary key for the demo data is ID. Let's make sure that we have a clean primary key.*
```{r}
check.demo <- demo %>%
  count(ID)
check.demo %>%
 	  filter(n > 1)
```



*1.2 tidy Job Prep GPA Data*
*1.2.1 Import and tidy Job Prep GPA Data file into a tibble named “gpa”.*
```{r}
gpa <- read_csv(here("Data","Job Prep GPA Data.csv"),
                 skip = 4, 
                 col_names = TRUE, 
                 col_types = list(ID = col_double(), 
                                  "School Code" = col_character(), 
                                  Age = col_character(),
                                  Gender = col_character()))

gpa
```


*1.2.2 Let's make sure that we have a clean key from the gpa file.*
```{r}
check.gpa <- gpa %>%
  count(ID)
check.gpa %>%
 	  filter(n > 1)
```


*1.2.3 It seems there have two students who have the same ID. Let's see what happened.*
```{r}
  subset(gpa, ID == 659)
# Did we typed the wrong ID number? We can look at other data to check for that.
```


*1.2.4 Let's compare gpa file with the demo file focusing on the student ID 659*
```{r}
subset(demo, ID == 659)
```


*1.2.5 We just can find the student ID 659 in School Code 4. Let's see if we could find if we type the wrong ID number in gpa file*
```{r}
subset (demo, "School Code" == 5, Age == 22, Gender = m)
# We could not find the student ID 659 in School code 5. We should remove this row faom gpa data. 
# If possilbe, I hope I could contact with the person who typed the data.
```


*1.2.6  Let's remove this row from gpa data. *
```{r}
gpa <- gpa %>%
  filter(!(ID == 659 & `School Code` == 5))
gpa
```


*1.2.7 Let's verify consistency between first and second semester. We could predict second semester GPA from first semester GPA*
```{r}
ggplot(data = gpa, mapping = aes(x = `S1 GPA`, y = `S2 GPA`)) +
  geom_point() +
   labs(x= "First Semester GPA", y = "Second Semester GPA", 
        title = "Scatterplot of first semester GPA and second semester GPA") + 
  stat_smooth()

```


*1.2.8 Let's average GPAs using a weighted average based on the number of credits taken in a semester.*
```{r}
gpa <- gpa %>%
  mutate(sygpa = (`S1 Credits`*`S1 GPA` + `S2 Credits`*`S2 GPA`)/(`S1 Credits` + `S2 Credits`) )
gpa
```

*1.3 Import and tidy Job Prep Survey file*
*1.3.1 import data and skip the title and empty rows*
```{r}
survey <- read_csv(here("Data","Job Prep Survey Results.csv"),
                 skip = 3, 
                 col_names = TRUE)
```

*1.3.2 account for reverse-scored items*
```{r}
survey[, c("Item 4", "Item 9", "Item 12", "Item 13")] <- 
  7 - survey[, c("Item 4", "Item 9", "Item 12", "Item 13")]

survey
```

*1.3.3 sum the items as an indicator of optimism about future job prospects*
```{r}
survey <- survey %>%
  mutate(optimism = rowSums(select(., 2:15)))

survey
```


*1.3.4 Let's make sure that we have a clean key from the survey file.*
```{r}
check.survey <- survey %>%
  count(ID)
check.survey %>%
 	  filter(n > 1)
```

*1.4 Let's combine the data. *

```{r}
data <-
  demo %>%
  left_join(gpa) %>%
  left_join(survey)
data
```


*1.5 Make the missing data explicit*
```{r}
data <- data[complete.cases(data),]
sum(table(data$ID)) == length(data$ID)

data

# 532/739 = 72%
# It is consistent with Mr Habernooth's letther, the response rate is 72%.
```
***


***
### Section #2
### Question 1: Is senior-year grade point average (GPA) related to the degree of optimism about future employment?
*2.1 Scatterplot of Senior-year GPA and optimism*
```{r}
ggplot(data = data, mapping = aes(x = sygpa, y = optimism)) +
  geom_point() +
   labs(x= "Senior-year GPA", y = "The degree of optimism about future employment", 
        title = "Scatterplot of Senior-year GPA and optimism") + 
  stat_smooth(method = "lm")
# From the scateerpoint, It seems senior-year grade point average (GPA) is not tightly related to the degree of optimism about future employment.
```


*2.2 Correlation of senior-year GPA and optimism*
```{r}
cor(data$sygpa,data$optimism)
# The correlation of senior-year GPA and optimism is in the lower level.
```


*2.3 Q-Q plot for residuals in the study of senior-year GPA and optimism *
```{r}
std_residual <- rstandard(model1)
qqnorm(std_residual)
qqline(std_residual)
# It seems there is the higher tails for Q-Q plot.
```
*2.3 Model summary*
```{r}
model1 <- lm(data$optimism ~ data$sygpa)
summary(model1)
# p = 0.642 > 0.01, we conclude that senior-year grade point average (GPA) is not significantly related to the degree of optimism about future employment.
```
***



***
### Section #3
### Question 2: Is the relationship of GPA and optimism consistent or different for different categories of disciplines?
*3.1 Boxplot of optimism scores by Disciplines*
```{r}
ggplot(data, aes(x = Discipline, y = optimism)) +
  geom_boxplot() +
  labs(x= "Discipline", y = "Optimism Scores", 
        title = "Optimism Scores by Disciplines")

```


*3.2 Summary data for optimism and discipline *
```{r}
tapply(data$optimism, data$Discipline, summary)
tapply(data$optimism, data$Discipline, sd)
```


*3.3 Boxplot of Senior-year GPA by Disciplines*
```{r}
ggplot(data, aes(x = Discipline, y = sygpa)) +
  geom_boxplot() +
  labs(x= "Discipline", y = "Senior-year GPA", 
        title = "Senior-year GPA by Disciplines")
```


*3.4 Summary data for Senior-year and discipline *
```{r}
tapply(data$sygpa, data$Discipline, summary)
tapply(data$sygpa, data$Discipline, sd)
```


*3.5 Scatterplot of Senior-year GPA and optimism by discipline*
```{r}
ggplot(data = data, mapping = aes(x = sygpa, y = optimism, col = Discipline)) +
  geom_point() +
   labs(x= "Senior-year GPA", y = "The degree of optimism about future employment", 
        title = "Scatterplot of Senior-year GPA and optimism") + 
  stat_smooth(method = "lm")
```

*3.6 check for interaction of Senior-year GPA and discipline*
```{r}
model2 <- data$optimism ~ data$sygpa*data$optimism
model2.out <- lm(model2)
anova(model2.out)
```

*3.7 Normal Q-Q plot* 
```{r}
model2.1 <- data$optimism ~ data$sygpa + data$Discipline
model2.1.out <- lm(model2.1)
plot(model2.1.out, 2)

```

*3.8 Summary of Senior-year GP,optimism and discipline*
```{r}
anova(model2.1.out)
# According the p<0.01 and the scatterplot of Senior-year GPA and optimism by discipline, we con conclude that the relationship of GPA and optimism is different for different categories of disciplines
```
***



***
### Section 4
### Question 3: Is the level of optimism different for different categories of disciplines?
*4.1 The boxplot of optimism by discipline*
```{r}
ggplot(data, aes(x = Discipline, y = optimism)) +
  geom_boxplot() +
  labs(x= "Discipline", y = "Optimism", 
        title = "Optimism by Discipline")

```


*4.2 Five-number summaries, means and standard deviations for optimism *
```{r}
tapply(data$optimism, data$Discipline, summary)
tapply(data$optimism, data$Discipline, sd)
```


*4.3 Normal Q-Q plot for residuals in the study of optimism and discipline*
```{r}
model3 <- data$optimism ~ data$Discipline
model3.out <- lm(model3)
ggplot(model3.out, aes(sample = rstandard(model3.out))) +
  stat_qq() +
  stat_qq_line() +
  labs(x = "Theoretical Quantiles",
       y = "Sample Quantiles",
      title = "Q-Q plot for residuals in the study of optimismand discipline")

```

*4.4 The model summary in the study of optimism and discipline*
```{r}
aov.model3 <- aov(data$optimism ~ data$Discipline)
aov.model3
summary(aov.model3)
# The null hypothesis of equal means can be rejected with p< 0.01.
# and conclude that the level of optimism is different for different categories of disciplines
```


*4.5 Tukey multiple comparisons of means*
```{r}
Tukey.results <- TukeyHSD(aov.model3)
Tukey.results
# From this table, we can conclude that most of pairs are different with 95% certain (p<0.01) except of Formal Sciences-Applied Sciences pair and Social Sciences-Natural Sciences pair,
```


